On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2015-11-23 MS.tex
Change main result to RGR exclusively analyses: Fig 3RGR, Fig4 all growth ms: rewrite the result part

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 232/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate (GR) has often been seen as a central element of its ecological strategy, and as ultimately linked to its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and GR, casting doubt on the capacity of traits to predict growth.

@@

\item Results were consistent with predictions. Specific leaf area was correlated with GR in seedling but not in adult plants. Correlations of GR with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with GR only in one plant stage category: seedlings and adults, respectively.

\item Synthesis: We show how trait-GR correlations change in a predictable way with plant size. Our understanding of plant strategies should shift away from attributing species as having slow or fast GR throughout their life, in favour of attributing growth trajectories.

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover (see section ``Theoretical expectations''). While recent empirical studies supported the idea of a variation in the relationships between traits and growth rates across plant sizes (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear. Moreover, the question whether the influence of traits on growth changes with plant size in a predictable and consistent way is very much open.

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3} a and d; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4} a-e). The correlation between growth and SLA shifted from strongly positive at seedling stage $\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{fig:fig3} a). The correlation between growth and SM shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult stage ($\bar{r}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{fig:fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between seedling and adult stages. The relationship between growth and SM in seedlings is expected to differ if absolute GR in mass is used instead of RGR mass. Here, for SM growth in mass was exclusively expressed in relative terms for seedlings (results shown in the Fig. \ref{fig:figA8} and Fig \ref{fig:figA4} f-j)

For WD, Hmax and Aarea, correlations with growth were also consistent with expectations, although with limitations arising from very unequal numbers of observations between growth stages for some traits. For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at intermediate and larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$). At sapling and seedling stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3} c), but also was consistent with no correlation (Fig \ref{fig:fig3} c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. We could not distinguish between these alternatives. Similarly for WD and for Aarea, results were consistent with negative (WD) and positive (Aarea) correlations with growth across all plant sizes (Fig. \ref{fig:fig3} b and e, model fit not improved when stage was incorporated). However, at the same time confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages.

@@

Given our basic conclusion that reported correlations responded to plant size in a manner consistent with expectations from theory, we assess now the coverage of the available data, to determine if biased sampling could have erroneously generated the above results.

In total, we found 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to  tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Number of species used for each correlation ranged from 2 to 300 species with average 31. Given this variety, it was not surprising to find wide ranges of $r$ reported for all traits. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}), and confounding with these factors is not a likely explanation of the major results reported.

The results reported above were obtained using the "ideal" dataset, and were thereby confined to correlations across species, and under non-stressed conditions ($n = 176$). The ``complete'' dataset ($n = 216$) also included correlations measured under stressed conditions, and spanning different plant stages between the growth and trait measurements. This ``complete'' dataset generated similar effect sizes to those reported above from the ``ideal'' dataset, and led to similar conclusions (Fig. \ref{fig:fig3} grey points). Differences primarily concerned the confidence intervals around particular mean effect sizes in Fig. \ref{fig:fig3}. These differences in CI may have arisen from a variety of causes. For instance, the correlation between growth and WD at seedling stage shifted from negative to non significant (i.e CI 95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct}; these correlations unlike others used values of WD from adult plants together with growth from seedlings. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 8 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{fig:figA5}).

@@

As far as we are aware, our study is the first to compare Aarea and growth correlations between small and large plants. In the growth model (Appendix \ref{sec:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. Meta-analyses showed a positive correlation for adult and seedling but not for saplings (n=2). In the seedling stage, where most of data were available, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig1.pdf}
\caption{Number of recorded correlations between trait and growth rate in our dataset, by size and trait. Traits are: Aarea (or Amass) = rate of CO$_{2}$ assimilation per unit leaf area (or leaf mass) = maximum photosynthetic rate, Hmax = asymptotic maximum height,  Ks = stem hydraulic conductance, LA = leaf area, LMR = leaf mass ratio, NARarea = net mass assimilation rate per unit of leaf area, Nmass (or Narea) = nitrogen content per unit of leaf mass (or leaf area), Pmass = phosphorus content per unit of leaf mass, SA/LA = sapwood area per leaf area, SLA = specific leaf area, thickness = leaf thickness, Vessel size = stem conduit size, vessel density = stem conduit density, WD = wood density. Results shown here are the ``complete'' dataset (see methods for details).}
\label{fig:fig1}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig3.pdf}
\caption{The relationship between growth rates and traits changes with plant stage and trait. Plots show the effect size (+ 95\%CI) for the average correlation across species, by plant stage. Effect size is a standardized measure of the deviation of correlation coefficient from zero (Fisher's z-transformed). Effects are significant if confidence intervals (95\%CI) for a given stage do not overlap with zero.  Trait names as in Fig. \ref{fig:fig1}. $n$ indicates number of correlations contributing to each effect. Likelihood Ratio Tests (LRT, for ``ideal'' dataset only) compared maximum-likelihood fits between a pooled model and one separated by stage. A significant p-value means the model including stage effect fitted the data better than the null model. Results shown are from both the ``complete'' (black)  and ``ideal'' (grey) datasets (see methods for details).}
\label{fig:fig3}
\end{figure}

>>>>

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 235/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate has often been seen as a central element of its ecological strategy, and as ultimately linked to its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and growth, casting doubt on the capacity of traits to predict growth.

@@

\item Results were consistent with predictions. Specific leaf area was correlated with relative growth rate (RGR) in seedling but not in adult plants. Correlations of RGR with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one plant stage category: seedlings and adults, respectively.

\item Synthesis: We show how trait-growth correlations change in a predictable way with plant size. Our understanding of plant strategies should shift away from attributing species as having slow or fast GR throughout their life, in favour of attributing growth trajectories.

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover (see section ``Theoretical expectations''). While recent empirical studies supported the idea of a variation in the relationships between traits and growth rates across plant sizes (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear.

@@

For SLA the correlation between RGR and the trait changed significantly with plant size in the manner expected. Model fit improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3} a; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4} a-e). The correlation between RGR and SLA shifted from strongly positive at seedling stage $\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{fig:fig3} a). The correlation coefficients recorded at sapling stage were intermediate between seedling and adult stages.

For WD, results were consistent with negative correlations with RGR across all plant sizes (Fig. \ref{fig:fig3} b, model fit not improved when stage was incorporated). Note that the confident intervals were not crossing zero for sapling and seedling stage despite small sample sizes (n= 4 and 2 respectively), meaning a good consistency in the coefficient reported for each stage across studies.

For both Hmax and seed mass, correlations with RGR were also consistent with expectations, although with limitations arising from very unequal numbers of observations between growth stages for some traits. Indeed, confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages, precluding strong conclusions. 
For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$). At sapling stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3} c, and confident intervals overlapping), but also was consistent with no correlation (Fig \ref{fig:fig3} c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. We could not distinguish between these alternatives. 
As expected for seed mass, the correlation with RGR shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult and sapling stages (Fig \ref{fig:fig3} c, confidence intervals overlapping zero). Hovever, the correlation between seed mass and RGR at adult and saling stages were also not significantly different from that observed at seedling stage (see log likelihood ratio tests and confident intervals overlapping in Fig. \ref{fig:fig3} c). Again we could not distinguish between these alternatives. But unlike Hmax, the number of coefficent reported for seed mass was almost similar between adult (n= 7) and seedling (n= 10) stages, suggesting that confident interval overalapping zero at adult stage were due to variation in coefficients of correlation reported rather than to a much smaller numbers of replicate.

For Aarea, results were consistent with a positive correlation with RGR at seedling stage. However, no conclusion can be drawn about a change along plant stages; only four correlation coefficients were reported for this trait, and three of them were measured at the seedling stage.

@@

Given our basic conclusion that reported correlations responded to plant size in a manner consistent with expectations from theory, we assess now the coverage of the available data, to determine if biased sampling could have erroneously generated the above results, or if additional data could reinforce some of our conclusions.

In total, we found only 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Number of species used for each correlation ranged from 2 to 300 species with average 31. Given this variety, it was not surprising to find wide ranges of $r$ reported for all traits. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}), and confounding with these factors is not a likely explanation of the major results reported.

The results reported above were obtained using an even more conservative dataset ($n = 98$), and were thereby confined to correlations with RGR, across at least 10 species, under non-stressed conditions, and controlling for the source of trait and growth (i.e. both had to be measured on the same stage). The ''complete'' dataset ($n = 216$) also included correlations measured with absolute growth rate, across two to 10 species, under stressed conditions, and spanning different plant stages between the growth and trait measurements. The release of criteria enlarged in a considerable extent the number of replicates. While some trait-stage combinations were particularly well studied with 59 correlation coefficient reported for SLA at seedling stage, other still showed weak support. Indeed, on the 13 traits-stage combinations studied here, six were supported by five or less correlations coefficient (i.e. WD-seedling, Hmax-sapling, Hmax-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling).

This ''complete'' dataset generated similar effect sizes to those reported above for SLA, WD and Hmax, and led to similar conclusions (Fig. \ref{fig:fig4}). While the correlation between growth and WD at seedling stage was still negative, the upper CI 95\% was very closed from zero (lower CI: -0.065) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct}; these correlations unlike others used values of WD from adult plants together with growth from seedlings.

For seed mass and Aarea, results from the ''complete'' dataset were even more consistent with expectations from the model. For seed mass, the correlation with RGR was undoubtly not significantly different from zero at adult stage (Fig \ref{fig:fig4}; CI for seedling and adult stages were not overlapping anymore, and model fit improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig4 d}. For Aarea, results were consistent with expectation of positive correlations with RGR across adult and seedling stages (Fig \ref{fig:fig4} e; CI not overlapping zero). The comparison of effect size between conservative dataset (Fig. \ref{fig:fig3 d}) and complete (Fig. \ref{fig:fig4 d} showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 8 correlations measured under low light conditions were added.

Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{fig:figA5}).

@@

As far as we are aware, our study is the first to compare Aarea and growth correlations between small and large plants. In the growth model (Appendix \ref{sec:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. Meta-analyses showed a positive correlation for adult and seedling but not for saplings (n=3). In the seedling stage, where most of data were available, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig1.pdf}
\caption{Number of recorded correlations between trait and growth rate in our dataset, by size and trait. Traits are: Aarea (or Amass) = rate of CO$_{2}$ assimilation per unit leaf area (or leaf mass) = maximum photosynthetic rate, Hmax = asymptotic maximum height,  Ks = stem hydraulic conductance, LA = leaf area, LMR = leaf mass ratio, NARarea = net mass assimilation rate per unit of leaf area, Nmass (or Narea) = nitrogen content per unit of leaf mass (or leaf area), Pmass = phosphorus content per unit of leaf mass, SA/LA = sapwood area per leaf area, SLA = specific leaf area, thickness = leaf thickness, Vessel size = stem conduit size, vessel density = stem conduit density, WD = wood density. Results shown here are the ''complete'' dataset (see methods for details).}
\label{fig:fig1}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig3.pdf}
\caption{The relationship between relative growth rates (RGR) and traits changes with plant stage and trait. Plots show the effect size (+ 95\%CI) for the average correlation across species, by plant stage. Effect size is a standardized measure of the deviation of correlation coefficient from zero (Fisher's z-transformed). Effects are significant if confidence intervals (95\%CI) for a given stage do not overlap with zero. Trait names as in Fig. \ref{fig:fig1}. $n$ indicates number of correlations contributing to each effect. Likelihood Ratio Tests (LRT) compared maximum-likelihood fits between a pooled model and one separated by stage. A significant p-value means the model including stage effect fitted the data better than the null model. Results shown are from correlation coefficients measured for RGR exclusively, with > 10 species, under unstressed conditions and for a controled trait source (see methods for details).}
\label{fig:fig3}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig4.pdf}
\caption{The relationships between growth rates and traits along plant stage are not overcasted by experimental conditions (i.e. growth measurement used, environemental conditions, trait source or number of species used to draw the correlation). Plots show the effect size (+ 95\%CI) for the average correlation across species, by plant stage. Effect size is a standardized measure of the deviation of correlation coefficient from zero (Fisher's z-transformed). Effects are significant if confidence intervals (95\%CI) for a given stage do not overlap with zero. Trait names as in Fig. \ref{fig:fig1}. $n$ indicates number of correlations contributing to each effect. Likelihood Ratio Tests (LRT) compared maximum-likelihood fits between a pooled model and one separated by stage. A significant p-value means the model including stage effect fitted the data better than the null model. Results shown are from the ''complete dataset'' (see methods for details).}
\label{fig:fig3}
\end{figure}
